---
title: "Microplastic flag degradation analysis"
output: html_document
date: "2024-11-30"
author: "Nadia Frontier"
---

# --------------------------------------------------------------

# *Environmental degradation and loss of polyester marker flags is a major source of localised plastic pollution in the Antarctic*

# --------------------------------------------------------------

## Clear R's brain
```{r}
rm(list = ls())
```

## Import data
```{r}
flags <- read.csv("flag_deployment_1_for_R.csv")
```


## Set factors

```{r}
flags$Type<-as.factor(flags$Type)
flags$ID<-as.factor(flags$ID)
```

## Install packages

```{r}
suppressWarnings(suppressPackageStartupMessages(library(car)))
suppressWarnings(suppressPackageStartupMessages(library(MASS)))
suppressWarnings(suppressPackageStartupMessages(library(nlme)))
suppressWarnings(suppressPackageStartupMessages(library(lme4)))
suppressWarnings(suppressPackageStartupMessages(library(multcomp)))
suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(library(emmeans)))
suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
suppressWarnings(suppressPackageStartupMessages(library(ggfortify)))

library(car)
library(MASS)
require (nlme)
require(lme4)
require(multcomp)
require(ggplot2)
require(emmeans)
require(dplyr)
require(ggfortify)
```


## Creating mean values for each material type

```{r}
mean<-tapply(flags$percent_lost, list(flags$Type), mean)
```

# Build model

```{r}
flagmodel <- lm(percent_lost~ Type, data= flags)
anova(flagmodel)
```
# Test residuals of the model

```{r}
hist(flagmodel$residuals) # slightly right skewed
```

# Plot residuals vs expected
```{r}
autoplot(flagmodel, smooth.colour = NA)
#The normal QQ plot looks quite good
#Not seeing much of a pattern in residuals vs fitted
```

# Post hoc testing

```{r}
emmeans(flagmodel, pairwise~Type)
```


# Using pipe to generate mean values within flag dataset

```{r}
mean1<- flags %>% 
group_by(Type) %>% 
summarise (percent_lost = mean (percent_lost))

print(mean1)
```

# Plot data

```{r}
lineplot<-ggplot(flags, aes(x = Type, y = percent_lost, colour = Type))+
  geom_boxplot(alpha=0.5)+
  geom_point(position= position_jitterdodge())+
  geom_point(data= mean1, size=7, shape =18, alpha=0.5)+
  ylab("Flag mass lost (%)")+
  xlab("Flag type")

lineplot+mytheme1
```

```{r}
ggsave(file= "Figure 5.pdf", width = 30, height = 21, scale=0.3)
```

